Healthcare Insights and Sociodemographic Analysis of New York: A Comprehensive Analysis of Patient Records in 2013 and 2021

### 2023-12-04 STAT405_FinalReport.qmd
Add files via upload

```{r, fig.width = 8, warning = FALSE}
#| echo: false
query <- dbSendQuery(dcon,"
SELECT *
FROM '2013data';
")
cleaned_data <- dbFetch(query, -1)
dbClearResult(query)
subsetted_data <- cleaned_data[cleaned_data$`HealthServiceArea` %in% c("New York City", "Western NY"), ]

@@

# Disparities in Charges and Costs

## Plot 3: Barplots of Average Total Charges by Race 2013 vs 2021

```{r, fig.height = 5, fig.width = 5}
#| echo: false
library(ggplot2)
library(patchwork)

@@

```{=tex}
\begin{center}
Figure 3: Barplots of Average Total Charges by Race in 2013 and 2021
\end{center}
```
In figure 3, the bar plot illustrates a compelling contrast in average total charges for hospital care among different racial groups. Notably, it highlights that individuals who identify as Black and those categorized under the "Other Races" group experience significantly higher average total charges compared to their White counterparts for their pregnancy charges. This disparity in healthcare costs sheds light on the existing inequalities within the healthcare system, prompting a critical examination of the factors contributing to these disparities and the potential need for targeted interventions to address the especially for marginalized populations of not just women on the basis of gender but also for race as well. We see that in 2021, post the COVID-19 pandemic, this disparity between average total charges of Black patients and White patients increased, showing that the pandemic potentially increased already present disparity in healthcare.

## Plot 4: Facet Density Plots of Total Cost per Age-Group Buckets

@@

Figure 4: Density Plots of Total Costs per Age-Groups

@@

## Plot 5: Pie Charts of Payment Type

@@

```{=tex}
\begin{center}
Figure 5: Pie Chart of Payment Type 
\end{center}
```
Our analysis delves into the intricacies of healthcare payments, as depicted by a comprehensive pie chart reflecting various payment types utilized by patients for hospital services. The chart initially reveals a dominance of Medicaid, Medicare, and private health insurance, indicating these as the predominant payment methods. Notably, Medicaid occupies a significant portion, underscoring its role in providing for individuals with limited financial resources. Medicare's substantial share reflects its critical support for the elderly and certain younger individuals with disabilities. The prominence of private health insurance highlights its importance in the American healthcare landscape, often associated with employment benefits.

@@

## Plot 6: Pie Charts of Payment Type by Race

@@

```{=tex}
\begin{center}
Figure 6: Pie Charts on Payment Type by Race
\end{center}
```
The pie charts shed light on a stark disparity in healthcare payment types across racial lines. They show that White patients are more likely to use Medicare, which often correlates with a more secure financial status and access to retirement benefits. Meanwhile, Black/African American individuals and those from other racial backgrounds have a higher reliance on Medicaid, indicating lower income levels and less access to private or employer-sponsored health insurance. These trends aren't just numbers; they reflect real-world inequalities where minority groups often have less access to quality healthcare. The data doesn't just highlight a disparity; it points to the need for policies that address the economic and racial factors that limit healthcare options for some groups. It's an important piece of the puzzle for creating a more equitable healthcare system where coverage and quality care are accessible to all, regardless of racial background or economic standing.

@@

### Plot 7: New York City Medicaid Heat Map

@@

```{=tex}
\begin{center}
Figure 7: Heat Map of Patients on Medicare in  New York City Boroughs
\end{center}
```
Building on the insights from Plot 6, which highlighted the distribution of healthcare payment types among different racial groups, our focus shifted to a more localized context---New York City. This targeted analysis allows us to dissect the intricate layers of Medicaid distribution by race within the city's diverse boroughs.

In Plot 7, the heatmap provides a stark visualization of racial disparities across NYC boroughs. The variance in color intensity, from light to dark red, denotes the number of Medicaid recipients, revealing a pronounced concentration in Kings (Brooklyn), particularly among Black/African American individuals. This contrast is particularly noteworthy when compared to Manhattan, where a higher count of White individuals is evident. The heatmap's gradient effectively communicates the disparity in Medicaid dependence between these two populous boroughs. Meanwhile, Richmond (Staten Island) shows markedly lower numbers for both racial groups, hinting at different underlying socioeconomic factors at play. The Bronx and Queens present a more balanced distribution, suggesting a more equitable spread of Medicaid coverage across racial lines in these areas.

### Plot 8: New York City Map Analysis

@@

```{=tex}
\begin{center}
Figure 8: Map of Patients on Medicare in  New York City Boroughs
\end{center}
```
Plot 8 deepens this exploration with a cartographic depiction that merges Medicaid data with the geographic layout of NYC. The boroughs are color-coded, and within each, circles of varying sizes represent the count of Medicaid recipients, with the color designating the race---blue for Black/African American and red for White. The visualization brings to light the sheer volume of Black/African American Medicaid beneficiaries in Brooklyn, as indicated by the dominance of the blue circles. In contrast, Manhattan's red circles point to a higher number of White individuals on Medicaid. This map is more than a visual aid; it encapsulates the demographic disparities in healthcare access and provides a spatial dimension to the quantitative data. For stakeholders, this map is a call to action, highlighting areas where healthcare policies and programs may need to be adjusted to address the specific needs of each borough's population, ensuring a more equitable healthcare system in one of the world's most diverse cities.

@@

@@

>>>>

```{r, fig.height = 4, fig.width = 8, warning = FALSE}
#| echo: false
query <- dbSendQuery(dcon,"
SELECT *
FROM '2013data';
")
cleaned_data <- dbFetch(query, -1)
dbClearResult(query)
subsetted_data <- cleaned_data[cleaned_data$`HealthServiceArea` %in% c("New York City", "Western NY"), ]

@@

# Disparities In Charges and Costs

## Plot 3: Linear Regression Model Regarding Charges and Costs

```{r, results='asis', warning=FALSE}
#| echo: false
query <- dbSendQuery(dcon,"
SELECT *
FROM '2013data';
")
cleaned_data <- dbFetch(query, -1)
dbClearResult(query)

# Renaming the column to 'Length of Stay' for the linear model
names(cleaned_data)[names(cleaned_data) == "LengthofStay"] <- "Length of Stay"

y <- cleaned_data$`TotalCharges`
length_of_stay <- strtoi(cleaned_data$`Length of Stay`)

# Set scipen to a large value to avoid scientific notation
options(scipen = 999999999999)

# Create a scatter plot
plot(length_of_stay, y, xlab = "Length of Stay", main = "Price vs Length of Stay", ylab = "Price")

# Fit the linear model
lmfs <- lm(y ~ length_of_stay)

# Add the regression line
abline(lmfs, col = "RED")

# Reset scipen to its default value after the plot
options(scipen = 0)
```

```{=tex}
\begin{center}
Figure 3: Length of Stay on Price Regression Dotplot
\end{center}
```
```{r, results='asis', message = FALSE, warning = FALSE, comment = FALSE}
#| echo: false
summary_lmfs <- summary(lmfs)
# Assuming lmfs is your linear model
summary_lmfs <- summary(lmfs)
r_squared <- summary_lmfs$r.squared

# Manually set the row names for the coefficients
rownames(summary_lmfs$coefficients) <- c("Intercept", "Length of Stay")

# Generate the table with xtable
table_lmfs <- xtable::xtable(summary_lmfs$coefficients, caption = "Linear Model Summary")
# Set column names for the table
colnames(table_lmfs) <- c("Estimate", "Std. Error", "t value", "Pr(>|t|)")
# Print the table w/ row names 
print(table_lmfs, include.rownames = TRUE, comment = FALSE)

```{r, results='asis', message = FALSE, warning = FALSE}
#| echo: false

summary_lmfs <- summary(lmfs)

# Extract R-squared value
r_squared <- summary_lmfs$r.squared

r_squared_df <- data.frame(`R^2` = r_squared)

table_r_squared <- xtable::xtable(r_squared_df, caption = "Model R-Squared")

print(table_r_squared, include.rownames = FALSE, 
      sanitize.text.function = function(x){x}, comment = FALSE)

In our analysis of patient data for the state of New York in 2013, we conducted a linear regression to explore the impact of the length of stay on the associated price. The results revealed a statistically significant and strong positive relationship between these variables. The coefficient for length of stay that is calcualte din days was estimated at 6140.73 with a standard error of 6.622 and a remarkably high t-value of 927.39. This suggests that as the length of stay increases, there is a substantial increase in the price, a relationship supported by a p-value close to zero. The R-squared value of 0.55 indicates that approximately 55% of the price variation can be explained by the length of stay, demonstrating its relevance as a predictor. These findings provide important insights for healthcare management and policymaking, highlighting the need to consider and manage the cost implications of varying lengths of stay for New York state patients in 2013.

However, this plot was the impetus to start seeing the interaction between total charges and costs and its relationship with other identifying factors of patients.

## Plot 4: Barplots of Average Total Charges by Race 2013 vs 2021

```{r, fig.height = 6, fig.width = 5}
#| echo: false
library(ggplot2)
library(patchwork)

@@

```{=tex}
\begin{center}
Figure 4: Barplots of Average Total Charges by Race in 2013 and 2021
\end{center}
```
In figure 4, the bar plot illustrates a compelling contrast in average total charges for hospital care among different racial groups. Notably, it highlights that individuals who identify as Black and those categorized under the "Other Races" group experience significantly higher average total charges compared to their White counterparts for their pregnancy charges. This disparity in healthcare costs sheds light on the existing inequalities within the healthcare system, prompting a critical examination of the factors contributing to these disparities and the potential need for targeted interventions to address the especially for marginalized populations of not just women on the basis of gender but also for race as well. We see that in 2021, post the COVID-19 pandemic, this disparity between average total charges of Black patients and White patients increased, showing that the pandemic potentially increased already present disparity in healthcare. We then wanted to start shifting our focus on looking at the age of patients.

## Plot 5: Facet Density Plots of Total Cost per Age-Group Buckets

@@

Figure 5: Density Plots of Total Costs per Age-Groups

@@

## Plot 6: Pie Charts of Payment Type

@@

```{=tex}
\begin{center}
Figure 6: Pie Chart of Payment Type 
\end{center}
```
Our analysis delves into the intricacies of healthcare payments, as depicted by a comprehensive pie chart reflecting various payment types utilized by patients for hospital services. The chart initially reveals a dominance of Medicaid, Medicare, and private health insurance, indicating these as the predominant payment methods. Notably, Medicaid occupies a significant portion, underscoring its role in providing for individuals with limited financial resources. Medicare's substantial share reflects its critical support for the elderly and certain younger individuals with disabilities. The prominence of private health insurance highlights its importance in the American healthcare landscape, often associated with employment benefits.

@@

## Plot 7: Pie Charts of Payment Type by Race

@@

```{=tex}
\begin{center}
Figure 7: Pie Charts on Payment Type by Race
\end{center}
```
The pie charts shed light on a stark disparity in healthcare payment types across racial lines. They show that White patients are more likely to use Medicare, which often correlates with a more secure financial status and access to retirement benefits. Meanwhile, Black/African American individuals and those from other racial backgrounds have a higher reliance on Medicaid, indicating lower income levels and less access to private or employer-sponsored health insurance. These trends aren't just numbers; they reflect real-world inequalities where minority groups often have less access to quality healthcare. The data doesn't just highlight a disparity; it points to the need for policies that address the economic and racial factors that limit healthcare options for some groups. It's an important piece of the puzzle for creating a more equitable healthcare system where coverage and quality care are accessible to all, regardless of racial background or economic standing.

@@

## Plot 8: New York City Medicaid Heat Map

@@

```{=tex}
\begin{center}
Figure 8: Heat Map of Patients on Medicare in  New York City Boroughs
\end{center}
```
Building on the insights from Plot 6, which highlighted the distribution of healthcare payment types among different racial groups, our focus shifted to a more localized context---New York City. This targeted analysis allows us to dissect the intricate layers of Medicaid distribution by race within the city's diverse boroughs.

In Plot 8, the heatmap provides a stark visualization of racial disparities across NYC boroughs. The variance in color intensity, from light to dark red, denotes the number of Medicaid recipients, revealing a pronounced concentration in Kings (Brooklyn), particularly among Black/African American individuals. This contrast is particularly noteworthy when compared to Manhattan, where a higher count of White individuals is evident. The heatmap's gradient effectively communicates the disparity in Medicaid dependence between these two populous boroughs. Meanwhile, Richmond (Staten Island) shows markedly lower numbers for both racial groups, hinting at different underlying socioeconomic factors at play. The Bronx and Queens present a more balanced distribution, suggesting a more equitable spread of Medicaid coverage across racial lines in these areas.

## Plot 9: New York City Map Analysis

@@

```{=tex}
\begin{center}
Figure 9: Map of Patients on Medicare in  New York City Boroughs
\end{center}
```
Plot 9 deepens this exploration with a cartographic depiction that merges Medicaid data with the geographic layout of NYC. The boroughs are color-coded, and within each, circles of varying sizes represent the count of Medicaid recipients, with the color designating the race---blue for Black/African American and red for White. The visualization brings to light the sheer volume of Black/African American Medicaid beneficiaries in Brooklyn, as indicated by the dominance of the blue circles. In contrast, Manhattan's red circles point to a higher number of White individuals on Medicaid. This map is more than a visual aid; it encapsulates the demographic disparities in healthcare access and provides a spatial dimension to the quantitative data. For stakeholders, this map is a call to action, highlighting areas where healthcare policies and programs may need to be adjusted to address the specific needs of each borough's population, ensuring a more equitable healthcare system in one of the world's most diverse cities.

@@

In our pursuit to enhance data transparency and support decision-making in healthcare, we introduce the Human Body Economic Diagram (HBED), a proposed analytical tool outlined in this report. The HBED represents a leap forward in medical data visualization, combining anatomical precision with economic analysis. It features an interactive silhouette of the human body, onto which data is mapped in an intuitive and accessible manner.

The HBED's core functionality lies in its ability to visually represent medical procedure frequencies and their associated costs on the human body. By using variably-sized circles, each corresponding to a different body part and scaled to indicate procedural frequency, and color-coding these circles to depict the average costs, the tool provides a quick, comparative view of the medical procedures landscape.

Complementing the diagram, a sophisticated dropdown menu will allow users to delve into the specifics of each anatomical region. Selecting a body part will reveal a detailed dashboard, presenting comprehensive metrics such as success rates of procedures, average recovery times, patient demographics, and regional cost variations. This granularity empowers users with the data needed to make informed decisions. The dropdown also includes Shiny elements to toggle between the year 2013 and 2021, and the body parts of head, chest, and torso.

Designed to be responsive and versatile, the HBED tool will be accessible across a range of devices, ensuring that whether a patient seeking cost transparency, a healthcare professional comparing procedural statistics, or a policymaker analyzing economic trends in healthcare, the information is readily available.

@@

```{r, fig.height= 3, fig.width=5, warning = FALSE, message = FALSE}
#| echo: false
library(shiny)
library(grid)
library(gridExtra)
library(scales)

query <- dbSendQuery(dcon,"
  SELECT *
  FROM '2021data';
  ")
  data_2021 <- dbFetch(query, -1)
  dbClearResult(query)
  query <- dbSendQuery(dcon,"
  SELECT *
  FROM '2013data';
  ")
  data_2013 <- dbFetch(query, -1)
  dbClearResult(query)

   # This function wraps your existing grid drawing code
  drawYourPlot <- function() {
    if(input$time == "2013"){
  cleaned_data <- data_2013
}
  if(input$time == "2021"){
  cleaned_data <- data_2021

}

  chest <- select(cleaned_data, APRMDCCode, TotalCharges, LengthofStay) %>% 
  filter(APRMDCCode %in% c(16, 5, 8, 4, 9, 21))

head <- select(cleaned_data, APRMDCCode, TotalCharges,LengthofStay) %>% 
  filter(APRMDCCode %in% c(20, 2, 1, 3, 10, 18, 19, 23))

stomach <- select(cleaned_data, APRMDCCode, TotalCharges, LengthofStay) %>% 
  filter(APRMDCCode %in% c(6, 13, 7, 11, 12, 17, 15, 24))

chest_count <- length(chest$APRMDCCode)
head_count <- length(head$APRMDCCode)
stomach_count <- length(stomach$APRMDCCode)
total_count <- length(cleaned_data$APRDRGCode)

draw_bed<- function(){
#bed frame
grid.draw(rectGrob(x = 0.80, y = 0.3, width = 0.375, height = 0.1, gp = gpar(fill = "brown3", col = "brown3", lwd = 2)))
#mattress
grid.draw(rectGrob(x = 0.80, y = 0.325, width = 0.35, height = 0.12, gp = gpar(fill = "gray", col = "gray", lwd = 2)))
#pillow
grid.draw(rectGrob(x = 0.675, y = 0.38, width = 0.125, height = 0.05, gp = gpar(fill = "gray", col = "darkgray", lwd = 2)))

#left headboard
grid.draw(rectGrob(x = 0.6, y = 0.35, width = 0.05, height = 0.35, gp = gpar(fill = "brown3", col = "brown3", lwd = 2)))

#right headboard
grid.draw(rectGrob(x = 0.975, y = 0.325, width = 0.05, height = 0.3, gp = gpar(fill = "brown3", col = "brown3", lwd = 2)))

#blanket
grid.draw(rectGrob(x = 0.85, y = 0.325, width = 0.2, height = 0.135, gp = gpar(fill = "lightpink", col = "pink2", lwd = 2)))

grid.text("Average Charge per Day", x = 0.8, y = 0.5, just = "centre", gp = gpar(fontsize = 10))

if(input$bodyPart == "Head"){
  grid.text(round(sum(head$TotalCharges)/sum(head$LengthofStay),2), x = 0.85, y = 0.33, just = "centre")
}
else if(input$bodyPart == "Chest"){
   grid.text(round(sum(chest$TotalCharges)/sum(chest$LengthofStay), 2), x = 0.85, y = 0.33, just = "centre")
}
else if(input$bodyPart == "Torso"){
   grid.text(round(sum(stomach$TotalCharges)/sum(stomach$LengthofStay), 2), x = 0.85, y = 0.33, just = "centre")
  }
}
head_gp <- gpar(fill = "white")
body_gp <- gpar(fill = "black")
arms_gp <- gpar(fill = "white")
legs_gp <- gpar(fill = "khaki")
if(input$bodyPart == "Total"){
grid.newpage()
color_palette <- colorRampPalette(c("green", "red"))
#Drawing the gradient key
grad_key_height <- 0.5
grad_key_width <- 0.05
grad_key_x <- 0.7
grad_key_y <- 0.5

draw_vertical_rectangle <- function(x, y, width, height, gp = gpar()) {
  grid.rect(x = x - width / 2, y = y - height / 2, width = width, height = height, gp = gp)
}

grid.rect(x = grad_key_x-.025, y = grad_key_y-0.005, width = grad_key_width, height = grad_key_height, gp = gpar(col = "black"))

colors <- color_palette(100)
for (i in 1:100) {
  rect_gp <- gpar(fill = colors[i], col = NA)
  y_pos <- grad_key_y - (grad_key_height / 2) + ((i - 1) / 100) * grad_key_height
  draw_vertical_rectangle(x = grad_key_x, y = y_pos, width = grad_key_width, height = grad_key_height / 100, gp = rect_gp)
}
# Add labels to the gradient key
grid.text("$", x = grad_key_x-0.015, y = grad_key_y - (grad_key_height / 2) - 0.02, just = "centre")
grid.text("$$$", x = grad_key_x-0.015, y = grad_key_y + (grad_key_height / 2) + 0.02, just = "centre")
grid.text("Costs", x = grad_key_x-0.015, y = grad_key_y + (grad_key_height / 2) + 0.1, just = "centre")

get_color <- function(count, max_count) {
  # Assuming 'color_palette' is a function or vector of colors that returns a color based on the index
  normalized_count <- count/max_count
  colors <- colorRampPalette(c("green", "yellow", "orange", "orange", "red"))(100)  # Blue for less, Red for more
  color_index <- max(1, ceiling(normalized_count*100))
  return(colors[color_index])
}
draw_circle <- function(x, y, radius, gp = gpar()) {
  grid.circle(x = x, y = y, r = radius, gp = gp)
}

max_group_cost <- max(mean(head$TotalCharges), mean(chest$TotalCharges), mean(stomach$TotalCharges))
head_count_gp <- gpar(fill = get_color(mean(head$TotalCharges), max_group_cost))
chest_count_gp <- gpar(fill = get_color(mean(chest$TotalCharges), max_group_cost))
stomach_count_gp <- gpar(fill = get_color(mean(stomach$TotalCharges), max_group_cost))

draw_vertical_rectangle(x = 0.3, y = 0.75, width = 0.1, height = 0.35, gp = arms_gp) # this is neck
draw_circle(x = 0.25, y = 0.78, radius = 0.11, gp = head_gp) # this is head
draw_vertical_rectangle(x = 0.35, y = 0.65, width = 0.2, height = 0.35, gp = body_gp) # this is body
draw_vertical_rectangle(x = 0.15, y = 0.65, width = 0.05, height = 0.35, gp = arms_gp) # this is left arm

draw_circle(0.25,0.78, (head_count/total_count)/10, head_count_gp)
draw_circle(0.30,0.60, (chest_count/total_count)/10, chest_count_gp)
draw_circle(0.25,0.43, (stomach_count/total_count)/10, stomach_count_gp)

draw_circle(0.85,.28, .03)
draw_circle(0.85,.36, .04)
draw_circle(0.85,.45, .045)
draw_circle(0.85,.555, .0525)
draw_circle(0.85,.68, .06)
grid.text("Less", x = 0.85, y = grad_key_y - (grad_key_height / 2) - 0.02, just = "centre")
grid.text("More", x = 0.85, y = grad_key_y + (grad_key_height / 2) + 0.02, just = "centre")
grid.text("# of Cases", x = 0.85, y = grad_key_y + (grad_key_height / 2) + 0.1, just = "centre")

draw_vertical_rectangle(x = 0.4, y = 0.65, width = 0.05, height = 0.35, gp = arms_gp) # this is right arm
draw_vertical_rectangle(x = 0.25, y = 0.35, width = 0.1, height = 0.3, gp = legs_gp) # this is left leg
draw_vertical_rectangle(x = 0.35, y = 0.35, width = 0.1, height = 0.3, gp = legs_gp) # this is right leg
grid.ellipse(x = 0.25, y = 0.88, size = 9, ar = 4, angle = 0, gp = gpar(fill = "black"))
}
    # Here you would add additional grid drawing code based on the input$bodyPart
    # For example:
    else if (input$bodyPart == "Head") {
      # Add grid elements specific to the head
      grid.newpage()

grid.circle(x = 0.25, y = 0.5, r = 0.35, gp = gpar(fill = "white", col = "black", lwd = 3))
grid.lines(x = c(0.15), y = c(0,0.19), 
           gp = gpar(col = "black", lwd = 3))
grid.lines(x = c(0.35), y = c(0,0.19), 
           gp = gpar(col = "black", lwd = 3))

colors <- colorRampPalette(c("red", "yellow","green"))(1000)
num_circles <- length(colors)

# CIRCLE
circle_radius <- 0.1
x_center <- 0.80
y_center <- 0.75

# GRADIETN FOR CIRCLE
for (i in 1:num_circles) {
  color <- colors[i]
  angle <- seq(0, pi, length.out = num_circles + 1)[i]
  x <- x_center + circle_radius * cos(angle) 
  y <- y_center + circle_radius * sin(angle) 
  #grid.draw()
  grid.circle(x = x, y = y, r = 50 / num_circles, gp = gpar(fill = color, col = color))
}
  #grid.rect(x = 0.3, y = 0.25, gp = gpar(fill = "white", col = "white"), width = 2, height = .5)
  # LINES for APR Code
# 4
grid.lines(x = c(0.8, 0.9), y = c(0.7, .8), 
           gp = gpar(col = "black", lwd = 3))

grid.ellipse(x = 0.25, y = 0.85, size = 30, ar = 4, angle = 0, gp = gpar(fill = "black"))
grid.text("APR Severity Level", x= 0.8, y = 0.95, just = "centre")
draw_bed()

    } else if (input$bodyPart == "Torso") {
      # Add grid elements specific to the torso
      grid.newpage()

      draw_circle <- function(x, y, radius, gp = gpar()) {
  grid.circle(x = x, y = y, r = radius, gp = gp)
}
draw_vertical_rectangle <- function(x, y, width, height, gp = gpar()) {
  grid.rect(x = x - width / 2, y = y - height / 2, width = width, height = height, gp = gp)
}
draw_vertical_rectangle(x = 0.45, y = 0.95, width = 0.4, height = 0.6, gp = body_gp) # this is body
draw_vertical_rectangle(x = 0.25, y = 0.35, width = 0.2, height = 0.4, gp = legs_gp) # this is left leg
draw_vertical_rectangle(x = 0.45, y = 0.35, width = 0.2, height = 0.4, gp = legs_gp) # this is right leg

colors <- colorRampPalette(c("red", "yellow","green"))(1000)
num_circles <- length(colors)

# CIRCLE
circle_radius <- 0.1
x_center <- 0.80
y_center <- 0.75

# GRADIETN FOR CIRCLE
for (i in 1:num_circles) { 
  color <- colors[i]
  angle <- seq(0, pi, length.out = num_circles + 1)[i]
  x <- x_center + circle_radius * cos(angle) 
  y <- y_center + circle_radius * sin(angle) 
  grid.circle(x = x, y = y, r = 50 / num_circles, gp = gpar(fill = color, col = color))
}
# LINES for APR Code

  # 2
grid.lines(x = c(0.75, .8), y = c(0.85, .7), 
           gp = gpar(col = "black", lwd = 3))

grid.text("APR Severity Level", x= 0.8, y = 0.95, just = "centre")
draw_bed()
    } else if (input$bodyPart == "Chest") {
      # Add grid elements specific to the chest
      grid.newpage()
draw_circle <- function(x, y, radius, gp = gpar()) {
  grid.circle(x = x, y = y, r = radius, gp = gp)
}
draw_vertical_rectangle <- function(x, y, width, height, gp = gpar()) {
  grid.rect(x = x - width / 2, y = y - height / 2, width = width, height = height, gp = gp)
}
draw_vertical_rectangle(x = 0.4, y = 0.55, width = 0.3, height = 0.6, gp = body_gp) # this is body
draw_vertical_rectangle(x = 0.10, y = 0.55, width = 0.05, height = 0.4, gp = arms_gp) # this is left arm
draw_vertical_rectangle(x = 0.45, y = 0.55, width = 0.05, height = 0.4, gp = arms_gp) # this is right arm

colors <- colorRampPalette(c("red", "yellow","green"))(1000)
num_circles <- length(colors)

# CIRCLE
circle_radius <- 0.1
x_center <- 0.80
y_center <- 0.75

# GRADIENT FOR CIRCLE
for (i in 1:num_circles) { 
  color <- colors[i]
  angle <- seq(0, pi, length.out = num_circles + 1)[i]
  x <- x_center + circle_radius * cos(angle) 
  y <- y_center + circle_radius * sin(angle) 
  grid.circle(x = x, y = y, r = 50 / num_circles, gp = gpar(fill = color, col = color))
}
# LINES for APR Code
#3
grid.lines(x = c(0.8, .85), y = c(0.7, .85), 
           gp = gpar(col = "black", lwd = 3))

grid.text("APR Severity Level", x= 0.8, y = 0.95, just = "centre")
draw_bed()
    }
  }

  input <- data.frame(time  = "2013",
                  bodyPart= "Total"
                  )

    drawYourPlot()

```{=tex}
\begin{center}
Figure 10: Initial Screen of HBED with Cases and Costs
\end{center}
```
In Figure 10 of the Human Body Economic Diagram (HBED) for 2013, colored circles on the human silhouette represent the average cost of medical procedures for the head, torso, and chest. The red circle on the head indicates high-cost procedures, likely due to specialized surgeries. The torso's orange circle suggests moderate costs, possibly from a range of surgical interventions, while the yellow circle on the chest points to lower costs, associated with more routine procedures. These color-coded circles quickly convey the cost disparities of medical procedures across different body regions, encapsulating the economic aspect of healthcare in 2013.

```{r, fig.height= 3, fig.width= 5, warning = FALSE, message = FALSE}
#| echo: false
library(shiny)
library(grid)
library(gridExtra)
library(scales)

query <- dbSendQuery(dcon,"
  SELECT *
  FROM '2021data';
  ")
  data_2021 <- dbFetch(query, -1)
  dbClearResult(query)
  query <- dbSendQuery(dcon,"
  SELECT *
  FROM '2013data';
  ")
  data_2013 <- dbFetch(query, -1)
  dbClearResult(query)

   # This function wraps your existing grid drawing code
  drawYourPlot <- function() {
    if(input$time == "2013"){
  cleaned_data <- data_2013
}
  if(input$time == "2021"){
  cleaned_data <- data_2021

}

  chest <- select(cleaned_data, APRMDCCode, TotalCharges, LengthofStay) %>% 
  filter(APRMDCCode %in% c(16, 5, 8, 4, 9, 21))

head <- select(cleaned_data, APRMDCCode, TotalCharges,LengthofStay) %>% 
  filter(APRMDCCode %in% c(20, 2, 1, 3, 10, 18, 19, 23))

stomach <- select(cleaned_data, APRMDCCode, TotalCharges, LengthofStay) %>% 
  filter(APRMDCCode %in% c(6, 13, 7, 11, 12, 17, 15, 24))

chest_count <- length(chest$APRMDCCode)
head_count <- length(head$APRMDCCode)
stomach_count <- length(stomach$APRMDCCode)
total_count <- length(cleaned_data$APRDRGCode)

draw_bed<- function(){
#bed frame
grid.draw(rectGrob(x = 0.80, y = 0.3, width = 0.375, height = 0.1, gp = gpar(fill = "brown3", col = "brown3", lwd = 2)))
#mattress
grid.draw(rectGrob(x = 0.80, y = 0.325, width = 0.35, height = 0.12, gp = gpar(fill = "gray", col = "gray", lwd = 2)))
#pillow
grid.draw(rectGrob(x = 0.675, y = 0.38, width = 0.125, height = 0.05, gp = gpar(fill = "gray", col = "darkgray", lwd = 2)))

#left headboard
grid.draw(rectGrob(x = 0.6, y = 0.35, width = 0.05, height = 0.35, gp = gpar(fill = "brown3", col = "brown3", lwd = 2)))

#right headboard
grid.draw(rectGrob(x = 0.975, y = 0.325, width = 0.05, height = 0.3, gp = gpar(fill = "brown3", col = "brown3", lwd = 2)))

#blanket
grid.draw(rectGrob(x = 0.85, y = 0.325, width = 0.2, height = 0.135, gp = gpar(fill = "lightpink", col = "pink2", lwd = 2)))

grid.text("Average Charge per Day", x = 0.8, y = 0.5, just = "centre", gp = gpar(fontsize = 10))

if(input$bodyPart == "Head"){
  grid.text(round(sum(head$TotalCharges)/sum(head$LengthofStay),2), x = 0.85, y = 0.33, just = "centre")
}
else if(input$bodyPart == "Chest"){
   grid.text(round(sum(chest$TotalCharges)/sum(chest$LengthofStay), 2), x = 0.85, y = 0.33, just = "centre")
}
else if(input$bodyPart == "Torso"){
   grid.text(round(sum(stomach$TotalCharges)/sum(stomach$LengthofStay), 2), x = 0.85, y = 0.33, just = "centre")
  }
}
head_gp <- gpar(fill = "white")
body_gp <- gpar(fill = "black")
arms_gp <- gpar(fill = "white")
legs_gp <- gpar(fill = "khaki")
if(input$bodyPart == "Total"){
grid.newpage()
color_palette <- colorRampPalette(c("green", "red"))
#Drawing the gradient key
grad_key_height <- 0.5
grad_key_width <- 0.05
grad_key_x <- 0.7
grad_key_y <- 0.5

draw_vertical_rectangle <- function(x, y, width, height, gp = gpar()) {
  grid.rect(x = x - width / 2, y = y - height / 2, width = width, height = height, gp = gp)
}

grid.rect(x = grad_key_x-.025, y = grad_key_y-0.005, width = grad_key_width, height = grad_key_height, gp = gpar(col = "black"))

colors <- color_palette(100)
for (i in 1:100) {
  rect_gp <- gpar(fill = colors[i], col = NA)
  y_pos <- grad_key_y - (grad_key_height / 2) + ((i - 1) / 100) * grad_key_height
  draw_vertical_rectangle(x = grad_key_x, y = y_pos, width = grad_key_width, height = grad_key_height / 100, gp = rect_gp)
}
# Add labels to the gradient key
grid.text("$", x = grad_key_x-0.015, y = grad_key_y - (grad_key_height / 2) - 0.02, just = "centre")
grid.text("$$$", x = grad_key_x-0.015, y = grad_key_y + (grad_key_height / 2) + 0.02, just = "centre")
grid.text("Costs", x = grad_key_x-0.015, y = grad_key_y + (grad_key_height / 2) + 0.1, just = "centre")

get_color <- function(count, max_count) {
  # Assuming 'color_palette' is a function or vector of colors that returns a color based on the index
  normalized_count <- count/max_count
  colors <- colorRampPalette(c("green", "yellow", "orange", "orange", "red"))(100)  # Blue for less, Red for more
  color_index <- max(1, ceiling(normalized_count*100))
  return(colors[color_index])
}
draw_circle <- function(x, y, radius, gp = gpar()) {
  grid.circle(x = x, y = y, r = radius, gp = gp)
}

max_group_cost <- max(mean(head$TotalCharges), mean(chest$TotalCharges), mean(stomach$TotalCharges))
head_count_gp <- gpar(fill = get_color(mean(head$TotalCharges), max_group_cost))
chest_count_gp <- gpar(fill = get_color(mean(chest$TotalCharges), max_group_cost))
stomach_count_gp <- gpar(fill = get_color(mean(stomach$TotalCharges), max_group_cost))

draw_vertical_rectangle(x = 0.3, y = 0.75, width = 0.1, height = 0.35, gp = arms_gp) # this is neck
draw_circle(x = 0.25, y = 0.78, radius = 0.11, gp = head_gp) # this is head
draw_vertical_rectangle(x = 0.35, y = 0.65, width = 0.2, height = 0.35, gp = body_gp) # this is body
draw_vertical_rectangle(x = 0.15, y = 0.65, width = 0.05, height = 0.35, gp = arms_gp) # this is left arm

draw_circle(0.25,0.78, (head_count/total_count)/10, head_count_gp)
draw_circle(0.30,0.60, (chest_count/total_count)/10, chest_count_gp)
draw_circle(0.25,0.43, (stomach_count/total_count)/10, stomach_count_gp)

draw_circle(0.85,.28, .03)
draw_circle(0.85,.36, .04)
draw_circle(0.85,.45, .045)
draw_circle(0.85,.555, .0525)
draw_circle(0.85,.68, .06)
grid.text("Less", x = 0.85, y = grad_key_y - (grad_key_height / 2) - 0.02, just = "centre")
grid.text("More", x = 0.85, y = grad_key_y + (grad_key_height / 2) + 0.02, just = "centre")
grid.text("# of Cases", x = 0.85, y = grad_key_y + (grad_key_height / 2) + 0.1, just = "centre")

draw_vertical_rectangle(x = 0.4, y = 0.65, width = 0.05, height = 0.35, gp = arms_gp) # this is right arm
draw_vertical_rectangle(x = 0.25, y = 0.35, width = 0.1, height = 0.3, gp = legs_gp) # this is left leg
draw_vertical_rectangle(x = 0.35, y = 0.35, width = 0.1, height = 0.3, gp = legs_gp) # this is right leg
grid.ellipse(x = 0.25, y = 0.88, size = 9, ar = 4, angle = 0, gp = gpar(fill = "black"))
}
    # Here you would add additional grid drawing code based on the input$bodyPart
    # For example:
    else if (input$bodyPart == "Head") {
      # Add grid elements specific to the head
      grid.newpage()

grid.circle(x = 0.25, y = 0.5, r = 0.35, gp = gpar(fill = "white", col = "black", lwd = 3))
grid.lines(x = c(0.15), y = c(0,0.19), 
           gp = gpar(col = "black", lwd = 3))
grid.lines(x = c(0.35), y = c(0,0.19), 
           gp = gpar(col = "black", lwd = 3))

colors <- colorRampPalette(c("red", "yellow","green"))(1000)
num_circles <- length(colors)

# CIRCLE
circle_radius <- 0.1
x_center <- 0.80
y_center <- 0.75
